---
title: 'Part 1: Time series wrangling & forecasting'
author: "Peyton Moore"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, PACKAGES}
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
```



```{r, READ IN DATA}
energy <- read_csv(here::here("data", "energy.csv"))


# some exploration

# ggplot(data = energy, aes(x = res_total, y = month))+
#   geom_point()

#bleck
```



```{r}
energy_ts <-  energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)

# now lets explore more 
# ggplot(data = energy_ts, aes(x = date, 
#                              y= res_total))+
#   geom_line()+
#   labs(y = "Residential energy consumption \n (Trillion BTU)")
# 
# # strong seasonality 
# # maybe an increasing trend
# 
# 
# ggplot(data = energy_ts, aes(x = date, y= ind_total))+
#   geom_line()+
#   labs(y = "Residential energy consumption \n (Trillion BTU)")
# # less clear, maybe seasonality... maybe cyclical? 


energy_ts %>% 
  gg_season(y = res_total)+
  theme_minimal()+
  labs(x = "month", 
       y = "residential energy consumption (trillion BTU)")

```

Key takeaways from season plots: 
- Highest residential energy consumption in December
- Another high point in July and August
- Low points in consumption in June and September
- There appears to be a trend in which energy consumption is increasing overtime 


```{r}
energy_ts %>%  
  gg_subseries(res_total)
```


```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total~ season()))

#View the components
components(dcmp) %>% autoplot()+
  theme_minimal()

```


```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```

Observations separated by 12 months are the most highly correlated with one another 
reflects the strong seasonality we've already seen 

```{r}
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# forecast using the model 10 years into the future: 
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

energy_forecast %>% 
  autoplot()


# or plot it added to the original data: 
energy_forecast %>% 
  autoplot(energy_ts)
```
### Assessing residuals 

```{r}
#Append the predicted values (and residuals) to original energy data 
energy_predicted <- broom::augment(energy_fit)

#Use View(energy_predicted) to see the resulting data frame


```

Now plot the actual energy values(res_total), and the predicted values (stored as .fitted) atop them:

```{r}
ggplot(data = energy_predicted)+
  geom_line(aes(x = date, y = res_total))+
  geom_line(aes(x = date, y = .fitted), color = "red")
```

lets explore residuals 
things to remember, residuals should be:
- uncorrelated
- centered at 0 
- normally distributed

```{r}
ggplot(data = energy_predicted, aes(x = .resid))+
  geom_histogram()
```
Looks normally distributed and centered at 0


