---
title: 'Part 1: Time series wrangling & forecasting'
author: "Peyton Moore"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, PACKAGES}
library(tidyverse)
library(tsibble)
library(feasts)
#library(fable)
```



```{r, READ IN DATA}
energy <- read_csv(here::here("data", "energy.csv"))


# some exploration

# ggplot(data = energy, aes(x = res_total, y = month))+
#   geom_point()

#bleck
```



```{r}
energy_ts <-  energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)

# now lets explore more 
# ggplot(data = energy_ts, aes(x = date, 
#                              y= res_total))+
#   geom_line()+
#   labs(y = "Residential energy consumption \n (Trillion BTU)")
# 
# # strong seasonality 
# # maybe an increasing trend
# 
# 
# ggplot(data = energy_ts, aes(x = date, y= ind_total))+
#   geom_line()+
#   labs(y = "Residential energy consumption \n (Trillion BTU)")
# # less clear, maybe seasonality... maybe cyclical? 


energy_ts %>% 
  gg_season(y = res_total)+
  theme_minimal()+
  labs(x = "month", 
       y = "residential energy consumption (trillion BTU)")

```

Key takeaways from season plots: 
- Highest residential energy consumption in December
- Another high point in July and August
- Low points in consumption in June and September
- There appears to be a trend in which energy consumption is increasing overtime 


```{r}
energy_ts %>%  
  gg_subseries(res_total)
```


```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total~ season()))

#View the components
components(dcmp) %>% autoplot()+
  theme_minimal()

```


```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot()
```

Observations separated by 12 months are the most highly correlated with one another 
reflects the strong seasonality we've already seen 

```{r}
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )
```



